---
title: "Expt2_behavioral_analyses"
author: "Anna Ivanova"
date: "3/16/2021"
output: html_document
---


```{r setup, include=FALSE}
rm(list = ls())    # clear workspace

knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
library(emmeans)
```

# Read and clean the data

```{r read data, echo=FALSE}
dat = read.csv('../Data/Expt2_data/study2.csv')

# clean the dataframe
dat = rename(dat, Dimension=Block.Type,Participant=participantID,
             CorrectAnswer=inCategory)
dat = dat %>%
  mutate(cBNT = scale(BNT)) %>%
  mutate(Group = recode(Group, 'Aphasia'='Aphasia',
                      'PD'='Parkinsons',
                      'NT'='Neurotypical')) %>%
  mutate(Accuracy = as.numeric(Accuracy)) %>% 
  mutate(HitRate = ifelse((Accuracy==1 & Response=='Y'), 1, 0)) %>%
  mutate(FalseAlarm = ifelse((Accuracy==0 & Response=='Y'), 1, 0)) %>%
  separate(Image.Name, c('Image.Name', NA), -4)

dat$FalseAlarm[is.na(dat$FalseAlarm)]=0

# Set up the contrasts and factor structure
# group - dummy coding with neurotypical as the reference level
# dimension - sum coding
dat$Group = factor(dat$Group, levels=c('Aphasia', 'Neurotypical', 'Parkinsons')) 
dat$Dimension = factor(dat$Dimension, levels=c('HD', 'LD'))
contrasts(dat$Dimension) = c(-0.5, 0.5)
colnames(attr(dat$Dimension, "contrasts")) = "LDGTHD"

dat$Participant = paste(dat$Participant, dat$Group, sep='_')
dat$Participant = factor(dat$Participant) 

# separate Hit Rate and False Alarm data
dat.HR = dat %>% filter(CorrectAnswer==1)
dat.FA = dat %>% filter(CorrectAnswer==0)
```

# Summary 

```{r data means, echo=FALSE}
# Dimension only
dat.summ.HR.dim = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Dimension) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA.dim = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Dimension) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ.RT.dim = dat %>% group_by(Group, Dimension, Participant) %>% 
  summarize(Response.Time = mean(Response.Time)) %>%
  group_by(Dimension) %>%
  summarize(meanRT = mean(Response.Time),
            sdRT = sd(Response.Time))

# Group only
dat.summ.HR.group = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Group) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA.group = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Group) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ.RT.group = dat %>% group_by(Group, Dimension, Participant) %>% 
  summarize(Response.Time = mean(Response.Time)) %>%
  group_by(Group) %>%
  summarize(meanRT = mean(Response.Time),
            sdRT = sd(Response.Time))


# by group and dimension 
dat.summ.HR = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ.RT = dat %>% group_by(Group, Dimension, Participant) %>% 
  summarize(Response.Time = mean(Response.Time)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanRT = mean(Response.Time),
            sdRT = sd(Response.Time))

# Case study - Aphasia 4
dat.summ.HR.A4= dat.HR %>% 
  filter(Participant=='4_Aphasia') %>%
  group_by(Dimension) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA.A4 = dat.FA %>%
  filter(Participant=='4_Aphasia') %>%
  group_by(Dimension) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ.RT.A4 = dat %>% 
  filter(Participant=='4_Aphasia') %>%
  group_by(Dimension) %>%
  summarize(meanRT = mean(Response.Time),
            sdRT = sd(Response.Time))


dat.summ = merge(dat.summ.HR, dat.summ.FA)
#write.csv(dat.summ, 'Tables/data_summary_Expt2.csv')
```

### By group

```{r, echo=FALSE}
m.HR = brm(HitRate ~ Group*Dimension + (1|Category) + (1|Participant),
                      family = bernoulli(),
                      data = dat.HR,cores = 4, iter = 4000, warmup= 1000)

summary(m.HR)

hypothesis(m.HR, 'GroupNeurotypical < 0')
hypothesis(m.HR, 'GroupParkinsons > 0')
hypothesis(m.HR, 'DimensionLDGTHD < 0')
hypothesis(m.HR,'GroupNeurotypical:DimensionLDGTHD < 0')
hypothesis(m.HR,'GroupParkinsons:DimensionLDGTHD < 0')
```

Follow-up comparisons:
```{r stats pairwise HR, echo=FALSE}
d = emmeans(m.HR, ~ Dimension) %>% 
    contrast("pairwise") 

emmeans(m.HR, ~ Group * Dimension) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")
```


#### With education as an additional covariate
```{r, echo=FALSE}
m.HR.edu = brm(HitRate ~ Group*Dimension + Education + (1|Category) + (1|Participant),
                      family = bernoulli(),
                      data = dat.HR,cores = 4, iter = 4000, warmup= 1000)

summary(m.HR.edu)
```

```{r, echo=FALSE}
#AIC(m.HR.edu, m.HR)
hypothesis(m.HR.edu, 'GroupNeurotypical < 0')
hypothesis(m.HR.edu, 'GroupParkinsons > 0')
hypothesis(m.HR.edu, 'DimensionLDGTHD < 0')
hypothesis(m.HR.edu,'GroupNeurotypical:DimensionLDGTHD < 0')
hypothesis(m.HR.edu,'GroupParkinsons:DimensionLDGTHD < 0')
hypothesis(m.HR.edu, 'Education > 0')
```


### By BNT

```{r, echo=FALSE}
m.HR.edu.BNT = brm(HitRate ~ cBNT*Dimension + Education + (1|Category) + (1|Participant), family = bernoulli(), data = dat.HR,cores = 4, iter = 4000, warmup= 1000)

summary(m.HR.edu.BNT)
```

### Model comparison

```{r, echo=FALSE}
hypothesis(m.HR.edu.BNT, 'cBNT > 0')
hypothesis(m.HR.edu.BNT, 'cBNT:DimensionLDGTHD > 0')
#modelcomp.HR = data.frame(
#  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
#  AIC=c(AIC(m.HR), AIC(m.HR.edu), AIC(m.HR.edu.BNT)))
# write.csv(modelcomp.HR, 'Tables/E1_ModelComparison_HitRate.csv')
```

## Plots

```{r, echo=FALSE}
dat2plot.HR = dat.HR %>%
  mutate(Group=factor(Group, levels=c('Neurotypical', 'Parkinsons', 'Aphasia'))) %>%
  group_by(Group,Dimension,Participant,BNT) %>%
  summarize(HitRate=mean(HitRate))

e1_hr = dat2plot.HR %>%
  ggplot(mapping=aes(x=Group, y=HitRate, fill=Dimension))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge', show.legend=FALSE)+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  coord_cartesian(ylim=c(0.5,1))+
  labs(y='Hit Rate', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical", "PD", "Aphasia"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=10))
e1_hr
```

```{r, echo=FALSE}
e1_hr_bnt = ggplot(data=dat2plot.HR)+
  geom_point(mapping=aes(x=BNT, y=HitRate, fill=Dimension),
            shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=HitRate, color=Dimension),
              alpha=0.2, size=1.5, show.legend=FALSE)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  coord_cartesian(ylim=c(0.72,1))+
  labs(y='Hit Rate')+
  theme_classic()+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=11),
        axis.text.y=element_text(size=11))

e1_hr_bnt
```

# False Alarms

## Stats

### By Group
```{r, echo=FALSE}
m.FA = brm(FalseAlarm ~ Group*Dimension + (1|Category) + (1|Participant), 
                      family = bernoulli(), data = dat.FA,cores = 4, iter = 4000, warmup= 1000) 
             #control = glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(m.FA)
```

####Follow-up comparisons
```{r, echo=FALSE}
d = emmeans(m.FA, ~ Dimension) %>% 
    contrast("pairwise") 

emmeans(m.FA, ~ Group * Dimension) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")

hypothesis(m.FA, 'GroupNeurotypical < 0')
hypothesis(m.FA, 'GroupParkinsons > 0')
hypothesis(m.FA, 'DimensionLDGTHD < 0')
hypothesis(m.FA,'GroupNeurotypical:DimensionLDGTHD < 0')
hypothesis(m.FA,'GroupParkinsons:DimensionLDGTHD < 0')
```

#### With education as an additional covariate
```{r, echo=FALSE}
m.FA.edu = brm(FalseAlarm ~ Group*Dimension + Education + (1|Category) + (1|Participant),
                      family = bernoulli(),
                      data = dat.FA,cores = 4, iter = 4000, warmup= 1000)

summary(m.FA.edu)
```

```{r, echo=FALSE}
#AIC(m.FA.edu, m.FA)
hypothesis(m.FA.edu, 'GroupNeurotypical < 0')
hypothesis(m.FA.edu, 'GroupParkinsons > 0')
hypothesis(m.FA.edu, 'DimensionLDGTHD < 0')
hypothesis(m.FA.edu,'GroupNeurotypical:DimensionLDGTHD < 0')
hypothesis(m.FA.edu,'GroupParkinsons:DimensionLDGTHD < 0')
```

### By BNT

```{r , echo=FALSE}
m.FA.edu = brm(FalseAlarm ~ Education + Group*Dimension + (1|Category) + (1|Participant), 
                      family = bernoulli(), 
                      data = dat.FA,cores = 4, iter = 4000, warmup= 1000)#,
#                 glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))

m.FA.edu.BNT = brm(FalseAlarm ~ Education + cBNT*Dimension + (1|Category) + (1|Participant), 
                      family = bernoulli(), 
                      data = dat.FA,cores = 4, iter = 4000, warmup= 1000)

summary(m.FA.edu.BNT)
```

### Model comparison
```{r, echo=FALSE}

hypothesis(m.FA.edu.BNT, 'cBNT < 0')
hypothesis(m.FA.edu.BNT, 'cBNT:DimensionLDGTHD < 0')

#modelcomp.FA = data.frame(
#  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
#  AIC=c(AIC(m.FA), AIC(m.FA.edu), AIC(m.FA.edu.BNT)))
#write.csv(modelcomp.FA, 'Tables/E1_ModelComparison_FalseAlarm.csv')
```

## Plots
```{r, echo=FALSE}
dat2plot.FA = dat.FA %>%
  mutate(Group=factor(Group, levels=c('Neurotypical', 'Parkinsons', 'Aphasia'))) %>%
  group_by(Group,Dimension,Participant,BNT) %>%
  summarize(FalseAlarm=mean(FalseAlarm))

e1_fa = ggplot(data=dat2plot.FA, mapping=aes(x=Group, y=FalseAlarm, fill=Dimension))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge', show.legend=FALSE)+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  coord_cartesian(ylim=c(0,0.11))+
  labs(y='False Alarm Rate', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical", "PD", "Aphasia"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  scale_y_continuous(breaks=c(0, 0.03, 0.06, 0.09))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=10))
e1_fa
```

```{r, echo=FALSE}
e1_fa_bnt = ggplot(data=dat2plot.FA)+
  geom_point(mapping=aes(x=BNT, y=FalseAlarm, fill=Dimension),
             shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=FalseAlarm, color=Dimension),
              alpha=0.2, size=1.5, show.legend=FALSE)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  labs(y='False Alarm Rate')+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=10),
        axis.text.y=element_text(size=10))

e1_fa_bnt
```





# RT

```{r , echo=FALSE}
dat.RT = read.csv('../Data/Expt1_data/study1_RT.csv')
dat.RT = rename(dat.RT, Dimension=category_type, Participant=participant,
                Category=category)
dat.RT = dat.RT %>%
  mutate(Group = recode(Group, 'NT'='Neurotypical', 'APH'='Aphasia',
                        'PD'='Parkinsons')) 
dat.RT$Participant = paste(dat.RT$Participant, dat.RT$Group, sep='_')

# Set up the contrasts and factor structure
dat.RT$Participant = factor(dat.RT$Participant)
# group - dummy coding with aphasia as the reference level
# dimension - sum coding
dat.RT$Group = factor(dat.RT$Group, levels=c('Aphasia', 'Neurotypical', 'Parkinsons')) 
dat.RT$Dimension = factor(dat.RT$Dimension, levels=c('HD', 'LD'))
contrasts(dat.RT$Dimension) = c(-0.5, 0.5)
colnames(attr(dat.RT$Dimension, "contrasts")) = "LD>HD"

dat.demographics = dat %>% 
  select(Participant, Education, cBNT, BNT) %>%
  distinct()

dat.RT = merge(dat.RT, dat.demographics)
```

## Stats

### Descriptive

```{r, echo=FALSE}
# Dimension only
dat.summ.RT.dim = dat.RT %>% group_by(Group, Dimension, Participant) %>% 
  summarize(RT = mean(TrialTime)) %>%
  group_by(Dimension) %>%
  summarize(meanRT = mean(RT),
            sdRT = sd(RT))

# Group only
dat.summ.RT.group = dat.RT %>% group_by(Group, Dimension, Participant) %>% 
  summarize(RT = mean(TrialTime)) %>%
  group_by(Group) %>%
  summarize(meanRT = mean(RT),
            sdRT = sd(RT))

# by group and dimension 
dat.summ.RT = dat.RT %>% group_by(Group, Dimension, Participant) %>% 
  summarize(RT = mean(TrialTime)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanRT = mean(RT),
            sdRT = sd(RT))
```

### By Group
```{r, echo=FALSE}
m.RT = brm(TrialTime ~ Group*Dimension + (1|Category) + (1|Participant), data = dat.RT, cores = 4, iter = 4000, warmup= 1000)
            #REML=FALSE)
summary(m.RT)
```

####Follow-up comparisons
```{r , echo=FALSE}
d = emmeans(m.RT, ~ Dimension) %>% 
    contrast("pairwise") 

emmeans(m.RT, ~ Group * Dimension) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")

hypothesis(m.RT, 'GroupNeurotypical < 0')
hypothesis(m.RT, 'GroupParkinsons > 0')
hypothesis(m.RT, 'DimensionLDGTHD < 0')
hypothesis(m.RT,'GroupNeurotypical:DimensionLDGTHD < 0')
hypothesis(m.RT,'GroupParkinsons:DimensionLDGTHD < 0')
```

### by BNT

```{r, echo=FALSE}
m.RT.edu = brm(TrialTime ~ Group*Dimension + Education + (1|Category) + (1|Participant), data = dat.RT,cores = 4, iter = 4000, warmup= 1000)

m.RT.edu.BNT = brm(TrialTime ~ cBNT*Dimension + Education + (1|Category) + (1|Participant), data = dat.RT, cores = 4, iter = 4000, warmup= 1000)

summary(m.RT.edu.BNT)
```

### Model comparison

```{r, echo=FALSE}
#modelcomp.RT = data.frame(
#  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
#  AIC=c(AIC(m.RT), AIC(m.RT.edu), AIC(m.RT.edu.BNT)))
#write.csv(modelcomp.RT, 'Tables/E1_ModelComparison_ResponseTime.csv')
```

## Plots

```{r, echo=FALSE}
dat2plot.RT = dat.RT %>%
  mutate(Group=factor(Group, levels=c('Neurotypical', 'Parkinsons', 'Aphasia'))) %>%
  group_by(Group,Dimension,Participant,BNT) %>%
  summarize(TrialTime=mean(TrialTime))

e1_rt = ggplot(data=dat2plot.RT, mapping=aes(x=Group, y=TrialTime, fill=Dimension))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge')+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  #coord_cartesian(ylim=c(0,0.1))+
  labs(y='Response Time (s)', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical", "PD", "Aphasia"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
                axis.text.x=element_text(size=14),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))
e1_rt
```


```{r, echo=FALSE}
e1_rt_bnt = ggplot(data=dat2plot.RT)+
  geom_point(mapping=aes(x=BNT, y=TrialTime, fill=Dimension),
            shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=TrialTime, color=Dimension),
              alpha=0.2, size=1.5)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  labs(y='Response Time (s)')+
  #coord_cartesian(ylim=c(0,65))+
  theme_classic()+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=11),
        axis.text.y=element_text(size=11),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))

e1_rt_bnt
```

# COMBINE

```{r, echo=FALSE}
(e1_hr |plot_spacer() |e1_fa |plot_spacer() |e1_rt |
   plot_layout(widths=c(1,0.1,1,0.1,1))) / 
  plot_spacer() / 
  (e1_hr_bnt |plot_spacer() |e1_fa_bnt |plot_spacer() |e1_rt_bnt |
     plot_layout(widths=c(1,0.1,1,0.1,1))) /
  plot_layout(heights=c(1,0.1,1))
ggsave('Plots/E1_results_wRT.png', width=36, height=18, units='cm')
```

# Accuracy

```{R echo=FALSE, message=FALSE, warning=FALSE}
m.Acc = brm(Accuracy ~ Group*Dimension + (1|Category) + (1|Participant),
                      family = bernoulli(),
                      data = dat.HR,cores = 4, iter = 4000, warmup= 1000)

summary(m.Acc)

hypothesis(m.Acc, 'GroupNeurotypical > 0')
hypothesis(m.Acc, 'GroupParkinsons > 0')
hypothesis(m.Acc, 'DimensionLDGTHD < 0')
hypothesis(m.Acc,'GroupNeurotypical:DimensionLDGTHD < 0')
hypothesis(m.Acc,'GroupParkinsons:DimensionLDGTHD < 0')
```

```{R, message=FALSE, warning=FALSE}
m.Acc.BNT = brm(HitRate ~ cBNT*Dimension + (1|Category) + (1|Participant),
               family = bernoulli(), data = dat.HR,
               cores = 4, iter = 4000, warmup= 1000)

hypothesis(m.Acc.BNT, 'cBNT > 0')
hypothesis(m.Acc.BNT, 'cBNT:DimensionLDGTHD > 0')

```