---
title: "Expt1_behavioral_analyses"
author: "Anna Ivanova"
date: "3/16/2021"
output: html_document
---


```{r setup, include=FALSE}
rm(list = ls())    # clear workspace

knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(afex)
library(emmeans)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
```

# Read and clean the data

```{r read data, echo=FALSE}

dat = read.csv('../data/Expt1_data/study1_responses.csv')

# Set up the contrasts and factor structure
# group - dummy coding with aphasia as the reference level
# dimension - sum coding
dat$Group = factor(dat$Group, levels=c('Aphasia', 'Neurotypical', 'Parkinsons')) 
dat$Dimension = factor(dat$Dimension, levels=c('HD', 'LD'))
contrasts(dat$Dimension) = c(-0.5, 0.5)
colnames(attr(dat$Dimension, "contrasts")) = "LD>HD"
dat$Participant = factor(dat$Participant)

# separate Hit Rate and False Alarm data
dat.HR = dat %>% filter(CorrectAnswer=='TRUE') 
dat.FA = dat %>% filter(CorrectAnswer=='FALSE')
```

# Summary 

```{r data means, echo=FALSE}
# Dimension only
dat.summ.HR.dim = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Dimension) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA.dim = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Dimension) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

# Group only
dat.summ.HR.group = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Group) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA.group = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Group) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

# by group and dimension 
dat.summ.HR = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ = merge(dat.summ.HR, dat.summ.FA)
write.csv(dat.summ, 'Tables/Supplementary/data_summary_Expt1_HR_FA.csv')
```

# Hit Rate

We are aiming to estimate the effect of dimensionality on the Hit Rate / False Alarm Rate / Response Time. We are treating category and participant as random effects. 

## Stats

### By group

```{r stats HR, echo=FALSE}
m.HR = glmer(HitRate ~ Group*Dimension + (1|Category) + (1|Participant),
                      family = binomial(),
                      data = dat.HR)

summary(m.HR)
```

Follow-up comparisons:
```{r stats pairwise HR, echo=FALSE}
d = emmeans(m.HR, ~ Dimension) %>% 
    contrast("pairwise") 

emmeans(m.HR, ~ Group * Dimension) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")
```


#### With education as an additional covariate
```{r stats HR edu, echo=FALSE}
m.HR.edu = glmer(HitRate ~ Group*Dimension + Education + (1|Category) + (1|Participant),
                      family = binomial(),
                      data = dat.HR)

summary(m.HR.edu)
```

```{r stats HR comparer, echo=FALSE}
anova(m.HR.edu, m.HR)
```


### By BNT

```{r stats HR, echo=FALSE}
m.HR.edu.BNT = glmer(HitRate ~ cBNT*Dimension + Education + (1|Category) + (1|Participant), family = binomial(), data = dat.HR)

summary(m.HR.edu.BNT)
```

### Model comparison

```{r model comparison HR, echo=FALSE}
modelcomp.HR = data.frame(
  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
  AIC=c(AIC(m.HR), AIC(m.HR.edu), AIC(m.HR.edu.BNT)))
write.csv(modelcomp.HR, 'Tables/Supplementary/E1_ModelComparison_HitRate.csv')
```

## Plots

```{r plot HR, echo=FALSE}
dat2plot.HR = dat.HR %>%
  mutate(Group=factor(Group, levels=c('Neurotypical','Aphasia', 'Parkinsons'))) %>%
  group_by(Group,Dimension,Participant,BNT) %>%
  summarize(HitRate=mean(HitRate))

e1_hr = dat2plot.HR %>%
  ggplot(mapping=aes(x=Group, y=HitRate, fill=Dimension))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge', show.legend=FALSE)+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  coord_cartesian(ylim=c(0.5,1))+
  labs(y='Hit Rate', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical","Aphasia", "PD"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=10))
e1_hr
```

```{r plot BNT, echo=FALSE}
e1_hr_bnt = ggplot(data=dat2plot.HR)+
  geom_point(mapping=aes(x=BNT, y=HitRate, fill=Dimension),
            shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=HitRate, color=Dimension),
              alpha=0.2, size=1.5, show.legend=FALSE)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  coord_cartesian(ylim=c(0.72,1))+
  labs(y='Hit Rate')+
  theme_classic()+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=11),
        axis.text.y=element_text(size=11))

e1_hr_bnt
```

# False Alarms

## Stats

### By Group
```{r stats FA, echo=FALSE}
m.FA = glmer(FalseAlarm ~ Dimension*Group + (1|Category) + (1|Participant), 
                      family = binomial(), data = dat.FA, 
             control = glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(m.FA)
```

####Follow-up comparisons
```{r stats pairwise FA, echo=FALSE}
d = emmeans(m.FA, ~ Dimension) %>% 
    contrast("pairwise") 

emmeans(m.FA, ~ Group * Dimension) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")
```

#### With education as an additional covariate
```{r stats FA edu, echo=FALSE}
m.FA.edu = glmer(FalseAlarm ~ Group*Dimension + Education + (1|Category) + (1|Participant),
                      family = binomial(),
                      data = dat.FA)

summary(m.FA.edu)
```

```{r stats HR comparer, echo=FALSE}
anova(m.FA.edu, m.FA)
```

### By BNT

```{r stats FA BNT, echo=FALSE}
m.FA.edu = glmer(FalseAlarm ~ Education + Dimension*Group + (1|Category) + (1|Participant), 
                      family = binomial(), 
                      data = dat.FA,
                 glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)))
m.FA.edu.BNT = glmer(FalseAlarm ~ Education + cBNT*Dimension + (1|Category) + (1|Participant), 
                      family = binomial(), 
                      data = dat.FA)

summary(m.FA.edu.BNT)
```

### Model comparison
```{r model comparison HR, echo=FALSE}
modelcomp.FA = data.frame(
  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
  AIC=c(AIC(m.FA), AIC(m.FA.edu), AIC(m.FA.edu.BNT)))
write.csv(modelcomp.FA, 'Tables/Supplementary/E1_ModelComparison_FalseAlarm.csv')
```

## Plots
```{r plot FalseAlarm, echo=FALSE}
dat2plot.FA = dat.FA %>%
  mutate(Group=factor(Group, levels=c('Neurotypical','Aphasia', 'Parkinsons'))) %>%
  group_by(Group,Dimension,Participant,BNT) %>%
  summarize(FalseAlarm=mean(FalseAlarm))

e1_fa = ggplot(data=dat2plot.FA, mapping=aes(x=Group, y=FalseAlarm, fill=Dimension))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge', show.legend=FALSE)+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  coord_cartesian(ylim=c(0,0.11))+
  labs(y='False Alarm Rate', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical","Aphasia", "PD"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  scale_y_continuous(breaks=c(0, 0.03, 0.06, 0.09))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=10))
e1_fa
```

```{r plot BNT, echo=FALSE}
e1_fa_bnt = ggplot(data=dat2plot.FA)+
  geom_point(mapping=aes(x=BNT, y=FalseAlarm, fill=Dimension),
             shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=FalseAlarm, color=Dimension),
              alpha=0.2, size=1.5, show.legend=FALSE)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  labs(y='False Alarm Rate')+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=10),
        axis.text.y=element_text(size=10))

e1_fa_bnt
```





# COMBINE

```{r combine plots, echo=FALSE}
(e1_hr |plot_spacer() |e1_fa |
   plot_layout(widths=c(1,0.1,1))) / 
  plot_spacer() / 
  (e1_hr_bnt |plot_spacer() |e1_fa_bnt |
     plot_layout(widths=c(1,0.1,1))) /
  plot_layout(heights=c(1,0.1,1))
ggsave('Plots/Supplementary/E1_results_HR_FA.png', width=25, height=20, units='cm')
```
