---
title: "Expt2_behavioral_analyses"
author: "Anna Ivanova"
date: "3/16/2021"
output: html_document
---


```{r setup, include=FALSE}
rm(list = ls())    # clear workspace

knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
library(emmeans)
```

# Read and clean the data

```{r read data, echo=FALSE}
dat = read.csv('Expt2_data/study2.csv')

# clean the dataframe
dat = rename(dat, Dimension=Block.Type,Participant=participantID,
             CorrectAnswer=inCategory)
dat = dat %>%
  mutate(cBNT = scale(BNT)) %>%
  mutate(Group = recode(Group, 'Aphasia'='Aphasia',
                      'PD'='Parkinsons',
                      'NT'='Neurotypical')) %>%
  mutate(Accuracy = as.numeric(Accuracy)) %>% 
  mutate(HitRate = ifelse((Accuracy==1 & Response=='Y'), 1, 0)) %>%
  mutate(FalseAlarm = ifelse((Accuracy==0 & Response=='Y'), 1, 0)) %>%
  separate(Image.Name, c('Image.Name', NA), -4)

dat$FalseAlarm[is.na(dat$FalseAlarm)]=0

# Set up the contrasts and factor structure
# group - dummy coding with neurotypical as the reference level
# dimension - sum coding
dat$Group = factor(dat$Group, levels=c('Aphasia', 'Neurotypical', 'Parkinsons')) 
dat$Dimension = factor(dat$Dimension, levels=c('HD', 'LD'))
contrasts(dat$Dimension) = c(-0.5, 0.5)
colnames(attr(dat$Dimension, "contrasts")) = "LD>HD"

dat$Participant = paste(dat$Participant, dat$Group, sep='_')
dat$Participant = factor(dat$Participant) 

# separate Hit Rate and False Alarm data
dat.HR = dat %>% filter(CorrectAnswer==1)
dat.FA = dat %>% filter(CorrectAnswer==0)
```

# Summary 

```{r data means, echo=FALSE}
# Dimension only
dat.summ.HR.dim = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Dimension) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA.dim = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Dimension) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ.RT.dim = dat %>% group_by(Group, Dimension, Participant) %>% 
  summarize(Response.Time = mean(Response.Time)) %>%
  group_by(Dimension) %>%
  summarize(meanRT = mean(Response.Time),
            sdRT = sd(Response.Time))

# Group only
dat.summ.HR.group = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Group) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA.group = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Group) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ.RT.group = dat %>% group_by(Group, Dimension, Participant) %>% 
  summarize(Response.Time = mean(Response.Time)) %>%
  group_by(Group) %>%
  summarize(meanRT = mean(Response.Time),
            sdRT = sd(Response.Time))


# by group and dimension 
dat.summ.HR = dat.HR %>% group_by(Group, Dimension, Participant) %>% 
  summarize(HitRate = mean(HitRate)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA = dat.FA %>% group_by(Group, Dimension, Participant) %>% 
  summarize(FalseAlarm = mean(FalseAlarm)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ.RT = dat %>% group_by(Group, Dimension, Participant) %>% 
  summarize(Response.Time = mean(Response.Time)) %>%
  group_by(Group, Dimension) %>%
  summarize(meanRT = mean(Response.Time),
            sdRT = sd(Response.Time))

# Case study - Aphasia 4
dat.summ.HR.A4= dat.HR %>% 
  filter(Participant=='4_Aphasia') %>%
  group_by(Dimension) %>%
  summarize(meanHitRate = mean(HitRate),
            sdHitRate = sd(HitRate))

dat.summ.FA.A4 = dat.FA %>%
  filter(Participant=='4_Aphasia') %>%
  group_by(Dimension) %>%
  summarize(meanFalseAlarm = mean(FalseAlarm),
            sdFalseAlarm = sd(FalseAlarm))

dat.summ.RT.A4 = dat %>% 
  filter(Participant=='4_Aphasia') %>%
  group_by(Dimension) %>%
  summarize(meanRT = mean(Response.Time),
            sdRT = sd(Response.Time))


dat.summ = merge(dat.summ.HR, dat.summ.FA)
#write.csv(dat.summ, 'Tables/data_summary_Expt2.csv')
```

# Hit Rate
## Stats
### By group

We are aiming to estimate the effect of dimensionality on the Hit Rate / False Alarm Rate / Response Time. We are treating category and participant as random effects. 

```{r stats HR, echo=FALSE}
m.HR = glmer(HitRate ~ Dimension*Group + (1|Category) + (1|Participant), 
                      family = binomial(), 
                      data = dat.HR)
summary(m.HR)
```

Follow-up comparisons:
```{r stats pairwise HR, echo=FALSE}
d = emmeans(m.HR, ~ Dimension) %>% 
    contrast("pairwise") 

emmeans(m.HR, ~ Group * Dimension) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")
```

### By BNT

```{r stats edu, echo=FALSE}
m.HR.edu = glmer(HitRate ~ Dimension*Group + Education + (1|Category) + (1|Participant), 
                      family = binomial(), 
                      data = dat.HR, control = glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

m.HR.edu.BNT = glmer(HitRate ~ Education + cBNT*Dimension + (1|Category) + (1|Participant), 
                      family = binomial(), data = dat.HR, 
                     control = glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(m.HR.edu.BNT)
```

### Model comparison
```{r model comparison HR, echo=FALSE}
modelcomp.HR = data.frame(
  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
  AIC=c(AIC(m.HR), AIC(m.HR.edu), AIC(m.HR.edu.BNT)))
write.csv(modelcomp.HR, 'Tables/E2_ModelComparison_HitRate.csv')
```

## Plots

```{r plot HR, echo=FALSE}
dat2plot.HR = dat.HR %>%
  mutate(Group=factor(Group, levels=c('Neurotypical', 'Aphasia', 'Parkinsons'))) %>%
  group_by(Group,Dimension,Participant,BNT) %>%
  summarize(HitRate=mean(HitRate))

e2_hr = ggplot(data=dat2plot.HR, mapping=aes(x=Group, y=HitRate, fill=Dimension))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge', show.legend=FALSE)+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  coord_cartesian(ylim=c(0.5,1))+
  labs(y='Hit Rate', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical", "Aphasia", "PD"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=10))

e2_hr 
```

```{r plot BNT, echo=FALSE}
e2_hr_bnt = ggplot(data=dat2plot.HR)+
  geom_point(mapping=aes(x=BNT, y=HitRate, fill=Dimension),
            shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=HitRate, color=Dimension),
              alpha=0.2, size=1.5, show.legend=FALSE)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  coord_cartesian(ylim=c(0.72,1))+
  labs(y='Hit Rate')+
  theme_classic()+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=11),
        axis.text.y=element_text(size=11))
e2_hr_bnt 
```


# False Alarms


## Stats

### By group
```{r stats FA, echo=FALSE}
m.FA = glmer(FalseAlarm ~ Dimension*Group + (1|Category) + (1|Participant), 
                      family = binomial(), 
                      data = dat.FA, control = glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(m.FA)
```

####Follow-up comparisons
```{r stats pairwise FA, echo=FALSE}
d = emmeans(m.FA, ~ Dimension) %>% 
    contrast("pairwise") 

emmeans(m.FA, ~ Group * Dimension) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")
```

### By BNT

```{r stats FA BNT, echo=FALSE}
m.FA.edu = glmer(FalseAlarm ~ Education + Dimension*Group + (1|Category) + (1|Participant), 
                      family = binomial(), 
                      data = dat.FA, control = glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
m.FA.edu.BNT = glmer(FalseAlarm ~ Education + cBNT*Dimension + (1|Category) + (1|Participant), 
                      family = binomial(), 
                      data = dat.FA, control = glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(m.FA.edu.BNT)
```

### Model comparison

```{r model comparison HR, echo=FALSE}
modelcomp.FA = data.frame(
  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
  AIC=c(AIC(m.FA), AIC(m.FA.edu), AIC(m.FA.edu.BNT)))
write.csv(modelcomp.FA, 'Tables/E2_ModelComparison_FalseAlarm.csv')
```

## Plots

```{r plot FalseAlarm, echo=FALSE}
dat2plot.FA = dat.FA %>%
  mutate(Group=factor(Group, levels=c('Neurotypical', 'Aphasia', 'Parkinsons'))) %>%
  group_by(Group,Dimension,Participant,BNT) %>%
  summarize(FalseAlarm=mean(FalseAlarm))

e2_fa = ggplot(data=dat2plot.FA, mapping=aes(x=Group, y=FalseAlarm, fill=Dimension))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge', show.legend=FALSE)+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.1, size=0.75,
               position=position_dodge(width=0.5))+
  #coord_cartesian(ylim=c(0,0.11))+
  labs(y='False Alarm Rate', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical", "Aphasia", "PD"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=10))
e2_fa
```

```{r plot BNT, echo=FALSE}
e2_fa_bnt = ggplot(data=dat2plot.FA)+
  geom_point(mapping=aes(x=BNT, y=FalseAlarm, fill=Dimension),
             shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=FalseAlarm, color=Dimension),
              alpha=0.2, size=1.5, show.legend=FALSE)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  labs(y='False Alarm Rate')+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=10),
        axis.text.y=element_text(size=10))
e2_fa_bnt
```

# RT

## Stats

### By Group
```{r stats RT, echo=FALSE}
m.RT = lmer(Response.Time ~ Dimension*Group + (1|Category) + (1|Participant), 
            data = dat, REML = FALSE)
summary(m.RT)
```

####Follow-up comparisons
```{r stats pairwise FA, echo=FALSE}
d = emmeans(m.RT, ~ Dimension) %>% 
    contrast("pairwise") 

emmeans(m.RT, ~ Group * Dimension) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")
```

### By BNT

```{r stats FA BNT, echo=FALSE}
m.RT.edu = lmer(Response.Time ~ Dimension*Group + Education + (1|Category) + (1|Participant), 
            data = dat, REML = FALSE)
m.RT.edu.BNT = lmer(Response.Time ~ Education + cBNT*Dimension + (1|Category) + (1|Participant), 
            data = dat, REML = FALSE)

summary(m.RT.edu.BNT)
```

### Model selection

```{r model comparison HR, echo=FALSE}
modelcomp.RT = data.frame(
  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
  AIC=c(AIC(m.RT), AIC(m.RT.edu), AIC(m.RT.edu.BNT)))
write.csv(modelcomp.RT, 'Tables/E2_ModelComparison_ResponseTime.csv')
```

## Plots

```{r plot FalseAlarm, echo=FALSE}
dat2plot.RT = dat %>%
  mutate(Group=factor(Group, levels=c('Neurotypical', 'Aphasia', 'Parkinsons'))) %>%
  group_by(Group,Dimension,Participant,BNT) %>%
  summarize(Response.Time=mean(Response.Time))

e2_rt = ggplot(data=dat2plot.RT, mapping=aes(x=Group, y=Response.Time, fill=Dimension))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge')+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  #coord_cartesian(ylim=c(0,0.1))+
  labs(y='Response Time (s)', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical", "Aphasia", "PD"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
                axis.text.x=element_text(size=14),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))
e2_rt
```

```{r plot BNT, echo=FALSE}
e2_rt_bnt = ggplot(data=dat2plot.RT)+
  geom_point(mapping=aes(x=BNT, y=Response.Time, fill=Dimension),
            shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=Response.Time, color=Dimension),
              alpha=0.2, size=1.5)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  labs(y='Response Time (s)')+
  #coord_cartesian(ylim=c(0,65))+
  theme_classic()+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=11),
        axis.text.y=element_text(size=11),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))

e2_rt_bnt
```


# COMBINE

```{r combine plots, echo=FALSE}
(e2_hr |plot_spacer() |e2_fa |plot_spacer() |e2_rt |
   plot_layout(widths=c(1,0.1,1,0.1,1))) / 
  plot_spacer() / 
  (e2_hr_bnt |plot_spacer() |e2_fa_bnt |plot_spacer() |e2_rt_bnt |
     plot_layout(widths=c(1,0.1,1,0.1,1))) /
  plot_layout(heights=c(1,0.1,1))
ggsave('Plots/E2_results_wRT.png', width=36, height=18, units='cm')
```

