---
title: "Expt1_behavioral_analyses"
author: "Anna Ivanova"
date: "2/26/2022"
output: html_document
---


```{r setup, include=FALSE}
rm(list = ls())    # clear workspace

knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(afex)
library(emmeans)
library(ggplot2)
library(dplyr)
library(tidyr)
library(patchwork)
```

# Read and clean the data

```{r read data, echo=FALSE}

dat = read.csv('../data/Expt1_data/study1_responses.csv')

# Set up the contrasts and factor structure
# group - dummy coding with aphasia as the reference level
# dimension - sum coding
dat$Group = factor(dat$Group, levels=c('Aphasia', 'Neurotypical', 'Parkinsons')) 
dat$Dimension = factor(dat$Dimension, levels=c('HD', 'LD'))
contrasts(dat$Dimension) = c(-0.5, 0.5)
colnames(attr(dat$Dimension, "contrasts")) = "LD>HD"
dat$Participant = factor(dat$Participant)
```

## Add alternative category type labels
```{r}
category_info = read.csv('../categories_recoded.csv')
dat = merge(dat, category_info)
contrasts(dat$sem_vs_perc) = c(-0.5, 0.5)
colnames(attr(dat$sem_vs_perc, "contrasts")) = "SEM>PERC"
```

# Summary 

```{r data means, echo=FALSE}
# sem_vs_perc only
dat.summ.dim = dat %>% group_by(Group, sem_vs_perc, Participant) %>% 
  summarize(Accuracy = mean(Accuracy)) %>%
  group_by(sem_vs_perc) %>%
  summarize(meanAccuracy = mean(Accuracy),
            sdAccuracy = sd(Accuracy))

# Group only
dat.summ.group = dat %>% group_by(Group, sem_vs_perc, Participant) %>% 
  summarize(Accuracy = mean(Accuracy)) %>%
  group_by(Group) %>%
  summarize(meanAccuracy = mean(Accuracy),
            sdAccuracy = sd(Accuracy))

# by group and sem_vs_perc 
dat.summ = dat %>% group_by(Group, sem_vs_perc, Participant) %>% 
  summarize(Accuracy = mean(Accuracy)) %>%
  group_by(Group, sem_vs_perc) %>%
  summarize(meanAccuracy = mean(Accuracy),
            sdAccuracy = sd(Accuracy))

# Trial inclusion
dat.num_trials = dat %>% 
  group_by(Participant, run_name) %>% 
  summarize(trials_included=length(trial_no)/20, 
            trials_excluded=32-trials_included)

write.csv(dat.summ, 'Tables/Supplementary/semperc/E1_data_summary.csv', row.names = FALSE)
write.csv(dat.num_trials, 'Tables/Supplementary/semperc/E1_num_trials.csv', row.names = FALSE)
```

# Accuracy

We are aiming to estimate the effect of category type (sem vs perc) on Accuracy & Response Time. We are treating category and participant as random intercepts. 

## Stats

### By group

```{r stats acc, echo=FALSE}
m.acc = glmer(Accuracy ~ Group*sem_vs_perc + (1|Category) + (1|Participant),
                      family = binomial(),
                      data = dat)

summary(m.acc)
```

#### With education as an additional covariate
```{r stats acc edu, echo=FALSE}
m.acc.edu = glmer(Accuracy ~ Group*sem_vs_perc + Education + (1|Category) + (1|Participant),
                      family = binomial(),
                      data = dat)

summary(m.acc.edu)
```

```{r stats acc compare, echo=FALSE}
anova(m.acc.edu, m.acc)
```


### By BNT

```{r stats acc, echo=FALSE}
m.acc.edu.BNT = glmer(Accuracy ~ cBNT*sem_vs_perc + Education + (1|Category) + (1|Participant), family = binomial(), data = dat)

summary(m.acc.edu.BNT)
```

### Model comparison

```{r model comparison acc, echo=FALSE}
modelcomp.acc = data.frame(
  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
  AIC=c(AIC(m.acc), AIC(m.acc.edu), AIC(m.acc.edu.BNT)))
write.csv(modelcomp.acc, 'Tables/Supplementary/semperc/E1_ModelComparison_Accuracy.csv')
```

## Plots

```{r plot acc, echo=FALSE}
dat2plot.acc = dat %>%
  mutate(Group=factor(Group, levels=c('Neurotypical', 'Aphasia', 'Parkinsons'))) %>%
  group_by(Group,sem_vs_perc,Participant,BNT) %>%
  summarize(Accuracy=mean(Accuracy))

e1_acc = dat2plot.acc %>%
  ggplot(mapping=aes(x=Group, y=Accuracy, fill=sem_vs_perc))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge', show.legend=FALSE)+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  coord_cartesian(ylim=c(0.5,1))+
  labs(y='Accuracy', x=NULL)+
  scale_x_discrete(labels=c("Neurotypical", "Aphasia", "PD"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=14),
        axis.text.y=element_text(size=10))
e1_acc
```

```{r plot BNT, echo=FALSE}
e1_acc_bnt = ggplot(data=dat2plot.acc)+
  geom_point(mapping=aes(x=BNT, y=Accuracy, fill=sem_vs_perc),
            shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=Accuracy, color=sem_vs_perc),
              alpha=0.2, size=1.5, show.legend=FALSE)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  coord_cartesian(ylim=c(0.5,1))+
  labs(y='Accuracy')+
  theme_classic()+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=11),
        axis.text.y=element_text(size=11))

e1_acc_bnt
```


# RT

```{r read data RT, echo=FALSE}
dat.RT = read.csv('../data/Expt1_data/study1_RT.csv')
dat.RT = rename(dat.RT, Dimension=category_type, Participant=participant,
                Category=category)
dat.RT = dat.RT %>%
  mutate(Group = recode(Group, 'NT'='Neurotypical', 'APH'='Aphasia',
                        'PD'='Parkinsons')) 
dat.RT$Participant = paste(dat.RT$Participant, dat.RT$Group, sep='_')

# Set up the contrasts and factor structure
dat.RT$Participant = factor(dat.RT$Participant)
# group - dummy coding with aphasia as the reference level
# dimension - sum coding
dat.RT$Group = factor(dat.RT$Group, levels=c('Aphasia', 'Neurotypical', 'Parkinsons')) 
dat.RT$Dimension = factor(dat.RT$Dimension, levels=c('HD', 'LD'))
contrasts(dat.RT$Dimension) = c(-0.5, 0.5)
colnames(attr(dat.RT$Dimension, "contrasts")) = "LD>HD"

dat.demographics = dat %>% 
  select(Participant, Education, cBNT, BNT) %>%
  distinct()

dat.RT = merge(dat.RT, dat.demographics)
```

## Add alternative category type labels
```{r}
dat.RT = merge(dat.RT, category_info)
contrasts(dat.RT$sem_vs_perc) = c(-0.5, 0.5)
colnames(attr(dat.RT$sem_vs_perc, "contrasts")) = "SEM>PERC"
```

## Stats

### Descriptive

```{r descr stats RT, echo=FALSE}
# sem_vs_perc only
dat.summ.RT.dim = dat.RT %>% group_by(Group, sem_vs_perc, Participant) %>% 
  summarize(RT = mean(TrialTime)) %>%
  group_by(sem_vs_perc) %>%
  summarize(meanRT = mean(RT),
            sdRT = sd(RT))

# Group only
dat.summ.RT.group = dat.RT %>% group_by(Group, sem_vs_perc, Participant) %>% 
  summarize(RT = mean(TrialTime)) %>%
  group_by(Group) %>%
  summarize(meanRT = mean(RT),
            sdRT = sd(RT))

# by group and sem_vs_perc 
dat.summ.RT = dat.RT %>% group_by(Group, sem_vs_perc, Participant) %>% 
  summarize(RT = mean(TrialTime)) %>%
  group_by(Group, sem_vs_perc) %>%
  summarize(meanRT = mean(RT),
            sdRT = sd(RT))
```

### By Group
```{r stats RT, echo=FALSE}
m.RT = lmer(TrialTime ~ sem_vs_perc*Group + (1|Category) + (1|Participant), data = dat.RT,
            REML=FALSE)
summary(m.RT)
```

#### With education as an additional covariate
```{r stats acc edu, echo=FALSE}
m.acc.edu = lmer(TrialTime ~ Group*sem_vs_perc + Education + (1|Category) + (1|Participant),
                      data = dat.RT, REML=FALSE)

summary(m.acc.edu)
```

####Follow-up comparisons
```{r stats pairwise FA, echo=FALSE}
d = emmeans(m.RT, ~ sem_vs_perc) %>% 
    contrast("pairwise") 

emmeans(m.RT, ~ Group * sem_vs_perc) %>% 
    contrast("pairwise", by = "Group") %>% 
    rbind(d) %>%
    summary(by=NULL, adjust="tukey")
```

### by BNT

```{r stats acc, echo=FALSE}
m.RT.edu = lmer(TrialTime ~ sem_vs_perc*Group + Education + (1|Category) + (1|Participant), data = dat.RT, REML=FALSE)

m.RT.edu.BNT = lmer(TrialTime ~ cBNT*sem_vs_perc + Education + (1|Category) + (1|Participant), data = dat.RT, REML=FALSE)

summary(m.RT.edu.BNT)
```

### Model comparison

```{r model comparison acc, echo=FALSE}
modelcomp.RT = data.frame(
  Model=c('Group x Dim', 'Group x Dim + Edu', 'BNT x Dim + Edu'),
  AIC=c(AIC(m.RT), AIC(m.RT.edu), AIC(m.RT.edu.BNT)))
write.csv(modelcomp.RT, 'Tables/Supplementary/semperc/E1_ModelComparison_ResponseTime.csv')
```

## Plots

```{r plot FalseAlarm, echo=FALSE}
dat2plot.RT = dat.RT %>%
  mutate(Group=factor(Group, levels=c('Neurotypical', 'Aphasia', 'Parkinsons'))) %>%
  group_by(Group,sem_vs_perc,Participant,BNT) %>%
  summarize(TrialTime=mean(TrialTime))

e1_rt = ggplot(data=dat2plot.RT, mapping=aes(x=Group, y=TrialTime, fill=sem_vs_perc))+
  stat_summary(geom = 'col', fun.y = 'mean', width=0.5, 
               position='dodge')+
  geom_point(position=position_jitterdodge(dodge.width=0.5, jitter.height=0, jitter.width=0.1),
             shape=21, size=1.5, alpha=0.5, show.legend = FALSE)+
  stat_summary(geom = 'errorbar', fun.data = 'mean_se', width=0.15, size=0.75,
               position=position_dodge(width=0.5))+
  #coord_cartesian(ylim=c(0,0.1))+
  labs(y='Response Time (s)', x=NULL, fill='Category type')+
  scale_x_discrete(labels=c("Neurotypical", "Aphasia", "PD"))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  theme_classic()+
  theme(axis.title.y=element_text(size=14),
                axis.text.x=element_text(size=14),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))
e1_rt
```


```{r plot BNT, echo=FALSE}
e1_rt_bnt = ggplot(data=dat2plot.RT)+
  geom_point(mapping=aes(x=BNT, y=TrialTime, fill=sem_vs_perc),
            shape=21, size=1.5, color='darkgray', show.legend=FALSE)+
  geom_smooth(method='lm', mapping=aes(x=BNT, y=TrialTime, color=sem_vs_perc),
              alpha=0.2, size=1.5)+
  scale_color_manual(values=c('#0A730A', '#98CD92'))+
  scale_fill_manual(values=c('#0A730A', '#98CD92'))+
  labs(y='Response Time (s)', color='Category type')+
  coord_cartesian(ylim=c(0,65))+
  theme_classic()+
  theme(axis.title.x=element_text(size=14),
        axis.title.y=element_text(size=14),
        axis.text.x=element_text(size=11),
        axis.text.y=element_text(size=11),
        legend.title = element_text(size=14),
        legend.text = element_text(size=12))

e1_rt_bnt
```

# COMBINE

```{r combine plots, echo=FALSE}
(e1_acc |plot_spacer() |e1_rt |
   plot_layout(widths=c(1,0.1,1))) / 
  plot_spacer() / 
  (e1_acc_bnt |plot_spacer() |e1_rt_bnt |
     plot_layout(widths=c(1,0.1,1))) /
  plot_layout(heights=c(1,0.1,1))
ggsave('Plots/Supplementary/semperc/E1_results_acc_RT.png', width=25, height=20, units='cm')
```

